---
title: "General Psychology Schedule Spring 2023"
author: "Brian Holt"
date: "`r Sys.Date()`"
output: 
  html_document:
   anchor_sections: TRUE
   css: style.css
   toc: true
   toc_float: true
   toc_depth: 6
   number_sections: no    # if true, section labels render
   config:
      toc:
        scroll_highlight: yes
        before: |
          <li><a href="./">General Psychology Spring 2023 schedule</a></li>`
        after: |
          <li><a href="https://github.com/rstudio/bookdown" target="blank">Published with bookdown</a></li>
      toolbar:
        position: fixed
  #  download: [ "epub"]  #Not sure how to set up for epub yet
      search: yes
bibliography: [General_references.bib]
biblio-style: apalike
link-citations: yes

description: "This is a minimal example of using the rmarkdown package for general psychology"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dagitty)
library(tidyverse)
library(ggdag)
library(tidygraph)
library(ggraph)


```




Introduce the concept of causality: Begin by explaining the difference between correlation and causation, and how DAGs can help us to understand causal relationships between variables.

Define DAGs: Define what a DAG is and how it can be used to represent causal relationships between variables. You can use examples such as a DAG showing the relationship between smoking and lung cancer, or the relationship between exercise and heart disease.





Basic structure of DAGs: Teach the basic structure of a DAG, including nodes (representing variables) and edges (representing causal relationships).




Explain the importance of directionality: Highlight the importance of directionality in a DAG, and how it represents the direction of causation.


# simple dagitty

```{r}
# Create a new DAG object
depression <- dagitty('dag{
  Genetics -> DepressionSymptoms
  LifeStressors -> DepressionSymptoms
  SocialSupport -> DepressionSymptoms
  }')

# Manually set the node coordinates
coordinates(depression) <- list(
  Genetics = c(0, 1),
  LifeStressors = c(0, 3),
  SocialSupport = c(1, 5),
  DepressionSymptoms = c(5, 3)
)

# View the DAG
plot(depression)
```


# using tidyverse and ggdag

```{r}

dagified <- dagify(DepressionSymptoms ~ Genetics,
                         DepressionSymptoms ~ LifeStressors,
                         DepressionSymptoms ~ SocialSupport,
                         exposure = c("LifeStressors",
                         "SocialSupport",
                         "Genetics"),
                         outcome = "DepressionSymptoms",
                   coords = list(
                     x = c(DepressionSymptoms = 5, 
                           Genetics = 1, 
                           LifeStressors = 1.5,
                           SocialSupport = 2.5),
                     y = c(DepressionSymptoms = 5, 
                           Genetics = 1, 
                           LifeStressors = 3,
                           SocialSupport=4.2)
))  


tidy_dagitty(dagified)



dagified %>% tidy_dagitty() %>% 
  ggplot(aes(x=x,y=y,xend = xend,
    yend = yend))+
    geom_dag_point(color='lightgreen') +
  geom_dag_edges() +
  geom_dag_text(color='blue',nudge_x = .5) +
  theme_dag()
  
ggdag( layout = "sugiyama") +
#  theme_dag() +
#  scale_size_continuous(range = c(10, 30)) +
  #geom_node_text(aes(label = name), size = 4, hjust = 1, vjust = 1) +
#  coord_flip() +
  theme(
    legend.position = "none",
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  )



```





Practice drawing DAGs: Provide opportunities for students to practice drawing DAGs, using examples from their own lives or from current events.

Discuss common misconceptions: Discuss common misconceptions about DAGs, such as assuming that all relationships are causal or assuming that all causal relationships are direct.

Analyze DAGs: Once students have a basic understanding of DAGs, you can move on to analyzing them. You can use R code to illustrate how to analyze DAGs, such as using the tidyverse package to create a DAG and ggdag to visualize it.



