---
title: "Student_goal_cleaning"
author: "Brian Holt"
date: "`r Sys.Date()`"
output: 
  html_document:
   css: style.css
   toc: true
   toc_float: true
   toc_depth: 6
   number_sections: false    # if true, section labels render
   config:
      toc:
        scroll_highlight: yes
        before: |
          <li><a href="./">Cleaning raw text</a></li>
        after: |
          <li><a href="https://github.com/rstudio/bookdown" target="blank">Published with bookdown</a></li>
      toolbar:
        position: fixed
  #    download: [ "epub"]  #Not sure how to set up for epub yet
      search: yes
bibliography: [General_references.bib]
biblio-style: apalike
link-citations: yes
description: "moving to content of raw text from students"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Read separate sheets from this link 



```{r, read sheet}
library(googlesheets4)
library(googledrive)
googledrive::drive_auth(email='bholt@northseattle.edu')


goals <-  'https://docs.google.com/spreadsheets/d/1nRM14XKDuq8XVBCB1Rdye3xwj9_oDqxnAXctTGyikgQ' 

#gs4_auth(goals)
# Count sheets  ?

classA<- read_sheet(ss=goal_sheet,sheet=1)
classB<- read_sheet(ss=goal_sheet,sheet=2)
classC<- read_sheet(ss=goal_sheet,sheet=3)
```
I'm expecting a list, but it could be a dataframe 

# in the text content, remove numbers larger than 4 digits. 

I am picking the number for because anything larger than that may start to resemble a student ID number, and although I'm sure there isn't any, I just want to make sure.

## So first count the number of times there are numbers larger than 4 digits

# Strip out gender reference in student name

Or at least put it into a new column


# Remove filler words like "and"
