---
title: "Survey Assignment"
author: "Brian Holt"
date: "`r Sys.Date()`"
output: revealjs::revealjs_presentation
# output: 
#   html_document:
#     toc: true
#     toc_float: true
#     toc_depth: 4
---

```{r setup, include=FALSE}
library(DiagrammeR)
```


# Intro

Essentially you are going to write a very short survey of 3 to 5 questions.  

There are a lot of steps we are going to skip for speed.

## Simple scientific method 



```{r flowchart, echo=FALSE}
DiagrammeR::grViz("
digraph flow {

  # left to right layout
  rankdir = LR;

  # node style
  node [shape = rectangle, style = rounded, fontsize = 10];

  subgraph cluster_science {
    label = 'Simple scientific method';
    Curious1           [label = 'Curious'];
    LibraryResearch    [label = 'Library research'];
    Hypothesis         [label = 'Generate hypothesis'];
    SetUpTest          [label = 'Set up test'];
    DataAnalysis       [label = 'Data analysis'];
    Conclude           [label = 'Conclude'];

    Curious1       -> LibraryResearch
    LibraryResearch -> Hypothesis
    Hypothesis     -> SetUpTest
    SetUpTest      -> DataAnalysis
    DataAnalysis   -> Conclude
  }


}
", width = "100%", height = "720px")
```



## Simple Survey Construction 

```{r flowchart survey, echo=FALSE}
DiagrammeR::grViz("
digraph flow {

  # left to right layout
  rankdir = LR;

  # node style
  node [shape = rectangle, style = rounded, fontsize = 10];

  subgraph cluster_survey {
    label = 'Survey development';
    Curious2        [label = 'Curious'];
    LibrarySurvey   [label = 'Library'];
    Brainstorm      [label = 'Brainstorm questions'];
    Pilot           [label = 'Pilot'];
    FinalQs         [label = 'Final selection of Qs'];
    Reliability     [label = 'Reliability'];
    Validity        [label = 'Validity'];

    Curious2     -> LibrarySurvey
    LibrarySurvey -> Brainstorm
    Brainstorm   -> Pilot
    Pilot        -> FinalQs
    FinalQs      -> Reliability
    Reliability  -> Validity
  }
}
", width = "100%", height = "720px")
```



# Due dates

There are 3 major parts: Generating questions, Piloting questions, then actually making the survey's available. 

* **Generate Questions Due: before class Thursday November 20th, Friday Nov 21st**

* **Pilot questions: Tuesday November 25th, Wednesday 26th**

* **Collect data by: Tuesday 12/2, Wednesday 12/3 **



# Explanations

* There is a worksheet in canvas that has a lot of information on what to do. 

* Survey's have basically 2 parts: the questions and then the demographics

* We are going to focus on generating questions at first. After Thanksgiving we can work on the data.

 **Generate Questions Due: Thurs 11/20, Friday 11/21**

 **Pilot questions: Tues 11/25, Wed 11/26**

 **Collect data by: Tues 12/2, Wed 12/3 **


